---
title: "Correlation and Regression"
output: learnr::tutorial
runtime: shiny_prerendered
tutorial:
  id: "regression"
  version: 0.5
---


```{r setup, include=FALSE}
library(learnr)
library(umdpsyc)
library(gradethis)
knitr::opts_chunk$set(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(
  exercise.timelimit = 60)
```

## Correlation

```{r}
data(exam_scores)
```

### Spearman's Rho


```{r}
cor(exam_scores$motivation,exam_scores$exam1,method = 'spearman')
```

### Pearson Correlation

```{r}
cor(exam_scores$sleep_week,exam_scores$exam1,method = 'pearson')
```
## Regression


## Packages

Many times, we want to use functions that other people have created and made available. We can do this easily using packages.

```{r, echo=TRUE}
library(openintro)
```

## Exercises

### Exercise 3

How would you find a linear regression model using scores on the first exam as the outcome and two predictor variables: minutes of sleep the week before and student motivation, including an interaction term? 


```{r exercise3, exercise=TRUE}

```

```{r exercise3-check}
grade_result(
  pass_if(~ identical(.result, summary(lm(exam1 ~ motivation*sleep_week, data = exam_scores))), "Good Job!"),
  fail_if(~ TRUE)
)
```

### Submitting work

```{r submission-setup}
submission_code <- function(UID, exercise = 'data-frames'){
  httr::sha1_hash('regression121',as.character(UID))
}
```

Generate your submission code by putting in your UID in the function below. For example, if your UID is `2`, then your code should look like `submission_code(UID = 2)`
```{r submission, exercise = TRUE}
submission_code(UID = )
```
