---
title: "Descriptive Statistics"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered

tutorial:
  id: "descriptive-statistics"
  version: 0.5
---


```{r setup, include=FALSE}
library(ggplot2)
library(learnr)
data(mtcars)
knitr::opts_chunk$set(echo = FALSE)
# learnr::tutorial_options(exercise.timelimit = 60)
```

## Descriptive Statistics

We use descriptive statistics to describe data.

## Measures of Center and Spread

Many of the functions for getting numerical summaries of center and spread are intuitive in R. 

|Function|Type of Plot|
|---|---|
|`mean(x)`|Finds the mean of a vector `x`|
| `median(x)` | Finds the median of a vector `x`.
| `sd(x)` | Finds the standard deviation of a vector `x` |
| `variance(x)` | Finds the variance of a vector `x` |
| `IQR(x)` | Finds the IQR of a vector `x`|
| `summary(x)` | Finds the min, 1st quartile, median, mean, 3rd quartile and max of a vector `x`, or of each column in a data frame `x`|


## Plotting with R

### Visualizing Numerical Variables


The basic functions for plotting numerical variables in `R` are shown in the table below. 

|Function|Type of Plot|
|---|---|
|`hist(x)`|histogram|
| `boxplot(x)` | boxplot |
| `plot(x,y)` | scatterplot |

For example, to get a simple histogram, we can use the following command.

```{r}
hist(mtcars$hp, breaks = 20) # The "breaks = 20" argument sets the number of bins
```


### Visualizing Categorical Variables

The base `R` graphics of categorical variables generally require you to give a table of the data as an argument. For example, let's make a basic barplot.

```{r}
barplot(table(mtcars$vs))
```

You can also use the `pie` function to create a pie chart, but we won't do that here. Remember, using a pie chart is highly discouraged.

To create a stacked bar chart, we can simply use a two-dimensional contingency table.

```{r}
barplot(table(mtcars$vs, mtcars$am))
```


## Exercises



### Submitting work

```{r submission-setup}
submission_code <- function(UID, exercise = 'data-frames'){
  print(round(UID * (213 / 214 + 102)*4+1))
}
```

Generate your submission code by putting in your UID in the function below. For example, if your UID is `2`, then your code should look like `submission_code(UID = 2)`
```{r submission, exercise = TRUE}
submission_code(UID = )
```
