---
title: "Descriptive Statistics for Numerical Variables"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered

tutorial:
  id: "numerical-descriptives"
  version: 0.5
---


```{r setup, include=FALSE}
library(ggplot2)
library(umdpsyc)
library(learnr)
library(gradethis)
data(mtcars)
knitr::opts_chunk$set(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
# learnr::tutorial_options(exercise.timelimit = 60)
```

## Descriptive Statistics

We use descriptive statistics to describe data.

## Measures of Center and Spread

Many of the functions for getting numerical summaries of center and spread are intuitive in R. 

|Function|Type of Plot|
|---|---|
|`mean(x)`|Finds the mean of a vector `x`|
| `median(x)` | Finds the median of a vector `x`.
| `sd(x)` | Finds the standard deviation of a vector `x` |
| `variance(x)` | Finds the variance of a vector `x` |
| `IQR(x)` | Finds the IQR of a vector `x`|
| `summary(x)` | Finds the min, 1st quartile, median, mean, 3rd quartile and max of a vector `x`, or of each column in a data frame `x`|


## Plotting with R

### Visualizing Numerical Variables

The basic functions for plotting numerical variables in `R` are shown in the table below. 

|Function|Type of Plot|
|---|---|
|`hist(x)`|histogram|
| `boxplot(x)` | boxplot |
| `plot(x,y)` | scatterplot |

For example, to get a simple histogram, we can use the following command.

```{r}
hist(mtcars$hp, breaks = 20) # The "breaks = 20" argument sets the number of bins
```

```{r}
boxplot(mtcars$hp)
```


## Exercises

Consider the `exam_scores` dataset.

```{r}
data(exam_scores)
head(exam_scores)
```

### Exercise 1

What is the overall mean of all exam scores? 

```{r exercise1, exercise=TRUE}

```

```{r exercise1-check}
grade_result(
  pass_if(~ identical(.result, mean(exam_scores$score, na.rm = TRUE)), "Good Job!"),
  fail_if(~ TRUE)
)
```


### Exercise 2

What is the overall mean of all exam scores? 

```{r exercise2, exercise=TRUE}

```

```{r exercise2-check}
grade_result(
  pass_if(~ identical(.result, mean(exam_scores$score, na.rm = TRUE)), "Good Job!"),
  fail_if(~ TRUE)
)
```

### Exercise 3

How would you find a histogram of the exam scores for all students who used a computer as their method of note-taking? 

**Hint:** This is a two step process. First, you need to subset, then you need to create the graph.

```{r exercise3, exercise=TRUE}

```

```{r exercise3-check}
grade_result(
  pass_if(~ identical(.result, hist(exam_scores[exam_scores$note_mode == 2,])), "Good Job!"),
  fail_if(~ TRUE)
)
```


### Submitting work

```{r submission-setup}
submission_code <- function(UID, exercise = 'data-frames'){
  httr::sha1_hash('numdesc',as.character(UID))
}
```

Generate your submission code by putting in your UID in the function below. For example, if your UID is `2`, then your code should look like `submission_code(UID = 2)`
```{r submission, exercise = TRUE}
submission_code(UID = )
```
