---
title: "T-tests"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
tutorial:
  id: "t-tests"
  version: 0.5
---


```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(umdpsyc)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(exercise.checker = gradethis::grade_learnr)
learnr::tutorial_options(
  exercise.timelimit = 60)

```

## Hypothesis Tests

```{r}
data("mtcars")
```

The steps to a hypothesis are:

1) State your hypotheses.

2) Set the criterion.

3) Collect sample data.

4) Calculate the test statistic.

5) Evaluate and interpret.

Most of these steps will be done outside of R, and in fact, they can generally only be done outside of R. However, R can be very useful for the computational part, step 4. 

In this tutorial, we will use R to calculate the test statistic as part of performing a t-test.

## Hypothesis Test Setup

$$H_0: \mu = 12000 $$
$$H_A: \mu \ne 12000$$


## Using `t.test()`

We can do the t-test by hand as we've shown, but `R` also has a convenient way of doing it using just the data. You do need to make sure you still go through all of the steps of performing a hypothesis test though --  the `t.test()` function will perform the test for you, but you need to make sure to provide all of the appropriate arguments.

Let's look at doing a simple two-sided test, with a null hypothesis that $\mu = 12000$. We provide the vector of data points, as well as an argument providing the null value. In this case, since the value is in thousands, we use `mu = 12`. 

```{r, echo = TRUE}
t.test(mtcars$mpg, mu = 12)
```

By default, this does a two-sided test. What if we wanted to do a one-sided test? We can use the `alternative` argument to change it.

```{r, echo = TRUE}
# These are how to do one-sided tests.
# Note: This is just to show that it's possible! Don't 
# change the sidedness after the fact like this!
t.test(mtcars$mpg, mu = 12, alternative = "greater")
t.test(mtcars$mpg, mu = 12, alternative = "less")
```

Note that this doesn't provide the conclusion of the test for you. You need to make the correct conclusion from these results. Basically, you just need to take the p-value and compare it to the alpha level like before. 

Lastly, notice that the output gives you information about the confidence interval. You can use the `t.test()` function to create confidence intervals as well. Just make sure you use the `conf.level` argument to specify the type of confidence interval you want.

```{r, echo = TRUE}
t.test(mtcars$mpg, conf.level = 0.99)
```

If you're just making a confidence interval, you can ignore the rest of the information. Here, I ran `t.test()` without anything for `mu` -- this gave me a test comparing it to 0, but I don't need that information since all I need is the confidence interval. 

## Exercises

### Exercise 1

Consider the `exam_scores` data that is included in the `umdpsyc` package. 

```{r exercises-setup}
data(exam_scores)
```

```{r exercise1-structure, echo = TRUE, exercise = TRUE}
str(exam_scores)
```

What is the mean?

### Exercise 2

How would you look at the top few rows of this dataset?

```{r exercise2, echo = TRUE, exercise = TRUE}

```

### Exercise 3

How would you subset the dataset to only include people who took notes by paper? 

```{r exercise3, echo = TRUE, exercise = TRUE}

```

```{r exercise3-check}
grade_result(
  pass_if(~ identical(.result, exam_scores[exam_scores$note_mode == 1,]), "Good job!"),
  fail_if(~ TRUE, "Not quite. Try again.")
)

```

### Exercise 4

```{r exercise4}
learnr::question("What would be the conclusion from this hypothesis test?",
         answer("Fail to reject the null hypothesis.", correct=FALSE),
         answer("Reject the null hypothesis.", correct = TRUE,  message = "Good job!"),
         allow_retry = TRUE,
         random_answer_order = TRUE
)
```

### Submitting work

```{r submission-setup}
submission_code <- function(UID, exercise = 'data-frames'){
  httr::sha1_hash('ttest123',as.character(UID))
}
```

Generate your submission code by putting in your UID in the function below. For example, if your UID is `2`, then your code should look like `submission_code(UID = 2)`
```{r submission, exercise = TRUE}
submission_code(UID = )
```
