---
title: "ANOVA"
output: learnr::tutorial
runtime: shiny_prerendered
tutorial:
  id: "anova"
  version: 0.5
---


```{r setup, include=FALSE}
library(learnr)
library(umdpsyc)
library(gradethis)
knitr::opts_chunk$set(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(
  exercise.timelimit = 60)
```

## ANOVA in R

We can use R to do ANOVA. 

This tutorial covers:
- Using R for doing ANOVA
- Interpreting R output for ANOVA
- Doing factorial ANOVA

## Using `aov`

Consider the `exam_scores` dataset. 

```{r}
data(exam_scores)
str(exam_scores)
```

```{r}
exam_anova <- aov(exam1 ~ factor(anxiety), data=exam_scores)
summary(exam_anova)
```


### ANOVA output



### 

## Factorial ANOVA



## Exercises

### Submitting work

```{r submission-setup}
submission_code <- function(UID, exercise = 'data-frames'){
  httr::sha1_hash('anova31',as.character(UID))
}
```

Generate your submission code by putting in your UID in the function below. For example, if your UID is `2`, then your code should look like `submission_code(UID = 2)`
```{r submission, exercise = TRUE}
submission_code(UID = )
```
