---
title: "ANOVA"
output: learnr::tutorial
runtime: shiny_prerendered
tutorial:
  id: "anova"
  version: 0.5
---


```{r setup, include=FALSE}
library(learnr)
library(umdpsyc)
library(gradethis)
knitr::opts_chunk$set(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
learnr::tutorial_options(
  exercise.timelimit = 60)
```

## ANOVA in R

We can use R to do ANOVA. 

This tutorial covers:
- Using R for doing ANOVA
- Interpreting R output for ANOVA
- Doing factorial ANOVA

## Using `aov`

Consider the `exam_scores` dataset. 

```{r}
data(exam_scores)
str(exam_scores)
```




## ANOVA output

Much of what you do in R will involve using **functions**. Let's take a look at a simple function.

```{r, echo=TRUE}
print('Hello')
```

Here, we are calling a function called `print`. This function takes in one argument, or input: `'Hello'`. The arguments go inside parentheses, and provide the input values. The function then takes that input and performs some sort of action. In this case, as you might have guessed, the `print` function prints out whatever is provided as input. 

### 

## Factorial ANOVA



## Exercises

### Submitting work

```{r submission-setup}
submission_code <- function(UID, exercise = 'data-frames'){
  httr::sha1_hash('anova31',as.character(UID))
}
```

Generate your submission code by putting in your UID in the function below. For example, if your UID is `2`, then your code should look like `submission_code(UID = 2)`
```{r submission, exercise = TRUE}
submission_code(UID = )
```
